Efficient Coalition Formation for Web Services

### 2013-03-10 papers/coalition/coalition.tex
Changes

Over the past years, web applications have become part of critical buisiness services. The increasing reliance on web-based applications has had significant implications for service providers and its users.

@@

TODO on community of web services \\

TODO Problems of \cite{Chun02user-centricperformance, mcknight00a, DBLP:journals/internet/BenatallahSD03, Khosravifar:2009:AIR:1586636.1586906, Rosario:2008:PQS:1512146.1512290} \\

TODO Problems of \cite {DBLP:conf/IEEEscc/LimTMB12} having to cop and optimizing for with all weak web services reduces the optimal result\\

TODO Problems of \cite {DBLP:conf/IEEEscc/KhosravifarABT11} not considering the current web services in coalition and startegies are only based on master web service and new web service gain\\

@@

We normalize and aggrigate each of these metrics to a real value between 0 and 1. Since tasks are not shared between web services and are distributed among all web serivices inside a community, one can expect tasks within a community to be performed by average quality of web service residing in that community. We use this average expected quality output, later in evaluating community $worth$ or $payoff$ function.

@@

\section{Problem Formulation and Modeling}

@@

\emph{Proposition: } The folowwing statement is equivalent with convex condition of equation \ref{eq:convex}; 
\begin{equation}\label{eq:convex_snow}
v(S \cup \left\{i\right\}) - v(S) \leq v (T \cup \left\{i\right\}) - v(T), \forall S \subseteq T \subseteq N \backslash \left\{i\right\}, \forall i \in N; 
\end{equation}

@@

In coalition-formation games, formation of the coalitions is the most important aspect of in the game. These solutions focus on maximizing social welfare. For any coalition structure $\pi$, let $v_{cs}(\pi)$ denote the total worth $\sum_{C \in \pi}{v(C)}$ which we call it social welfare. The problems in this area deal with finding the maximum value for social welfare over all the possible coalition structures $\pi$. There are $centralized$ algorithms for this end, however these approaches are generally NP-complete. The reason is that number of all possible partitions of a set $N$ grows exponentially with the number of players in $n$, and the centralized algorithms need to iterate through all these partitions. 
%These algorithms \cite{DBLP:conf/ijcai/GrecoMPS11, DBLP:conf/ijcai/RahwanMJ11, RePEc:wpa:wuwpga:0110001} are centralized algorithms, where all the complexity. However these algorithms are more intractable than Core stable solutions and practical with some constraints in practice\cite{RePEc:wpa:wuwpga:0110001}. 
In our application, we prefer not having a centralized decision maker dictating all web services their decisions. Becuase of this plus the complexity issues we would prefer a distributed algorithm where each community master can be a decision maker. The aim is to find low complexity and distributed algorithms for forming coalitions\cite{DBLP:journals/igtr/AptW09,Dieckmann02dynamiccoalition,ray2007game}. The distributed merge-and-split algorithm in \cite{DBLP:journals/igtr/AptW09} suits our application very well. It keeps splitting and merging coalitions where it is preferred by all players inside those coalitions. 
This merge-and-split algorithm is designed to be adoptable for different problems. It two functions to be specified, first a preference relation or comparison relation for comparing two collections of different coalition partitions of same players and second function is to assess the stability condition of a partition. Having two partition sets of a subset of our players, namely $P = {P_1,...,P_k}$ and $Q = {Q_1,...N_l}$ one example would be to use the social welfare comarison $\sum^k_{i=1}v(P_i) > \sum^k_{i=1}v(N_l)$. For our scenario we used $Pareto order$ comparison which is an individual-value order appropriate for our self-interest agents (web services). In Pareto order, an allocation or partition $P$ is preferred over another $Q$ if at least one player gets more payoff in new allocation without other players getting hurt, or other players at least same payoff. ($p_i \geq q_i$ with at least one element $p_i > q_i$). As new web services are discovered and get ready to join communities, this algorithm keeps merging and splitting partitions based on the preference or comparison function provides and will stop as the partitions satisfy the stability condition or no further merging and splitting is found satisfying the comparison function. More details of this algorithm and analyses of generic solutions on coalition formation games are described in \cite{DBLP:journals/igtr/AptW09}.

@@

\section{Experimental Results}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[scale=0.61]{task_done_opt.eps}
\includegraphics[scale=0.61]{task_qos_opt.eps}
\caption{Part (a): Cumulative number of tasks succesfully done. Part
(b): Average QoS of tasks performed.} \label{performanceall}
\end{figure}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[scale=0.6]{least_core.eps}
\caption{Smallest value of $\epsilon$ such that the $\epsilon-core$ is non-empty} \label{f_leastcore}
\end{figure}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[scale=0.61]{s2_task_done.eps}
\includegraphics[scale=0.61]{s2_task_qos.eps}
\caption{Part (a): Cumulative number of tasks succesfully done. Part
(b): Average QoS of tasks performed.} \label{performancemany}
\end{figure}

@@

\section{Conclusion}

>>>>

Over the past years, web applications have become part of critical buisiness services. The increasing reliance on web-based applications has significant implications for service providers and its users. Web services are methods of communication between electronic devices over the internet. Web service as a highly available software function acessable at a network address over the web or the cloud usually provides a set of \emph{stateless} functions.

@@

The need for highly available and highly responsive services online has called for collaborative communities of these loosely-coupled business services. The idea of web service communities and the way they are engineered to collaborate was proposed in \cite{DBLP:journals/ijebr/MaamarSTBB09, DBLP:journals/internet/BenatallahSD03, Rosario:2008:PQS:1512146.1512290} where web services with similar functionality are grouped together and users send their requests to communities assuring higher service availability and performance than individual web services. These communities are highly available also more resilience to failure since there is no single point of failure in their architecture.

Most of the recent work on communities of services are either user-centric and focus user satisfaction \cite{Chun02user-centricperformance} or there are many works in distributed, grid, cluster and cloud computing which are system-centric and focus on system throughput and performance and utilization of the system. However in real world environments both users and services are selfish and self-interested agents, aiming to maximize their own profit. In real world environments, both these parties will participate and collaborate as long as its economically efficient for them.

Recently, \cite{DBLP:conf/IEEEscc/LimTMB12, DBLP:conf/IEEEscc/KhosravifarABT11} have proposed mechanisms in which try to maximize gain for both parties. A two-player non-cooperative game theory method between web services and community manager was proposed in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. The strategies for a web service when facing a new communities, is joining or staying in their community and the set of strategies for communities are allowing to resusing the the web service proposal. Both parties based on their capacity, market share and reputation will have different set of utilities over the strategy profiles of the game. The problem with this method is the game is between community master and the new web service, and all other web services already residing within the community are ignored. Those web services can lose utility and gain having to collabrate with new webserices, they may deviate and join other comuunities if they are unsatisfied. In \cite{DBLP:conf/IEEEscc/LimTMB12} a 3-Way satisfaction selection approach has been proposed. In thier approach, they proposed a web service selection process for community masters catering for the effciency of 3 parties of users, web services and communities. In this work, they show how the gain of these parties are coupled together and they suggest a linear optimization process for optimizing the needs of the 3 parties. However, the optimization problem in this solution tends to optimize some parameters considering all web serivces even if they are weak. There are no clear thresholds for accepting or rejecting new web services. The optimization result for example could suggest web services already residing within the community to increase or decrease their capacity to adjust with weakness of other parties in the system. Hoever, a strong performing web service could deviate anytime it finds itself unsatisfied within the community instead of adjusting its service parameters.

In this paper, we proposed a cooperative game theory based model for aggregation of web services within communities. The solution concepts of cooperative game theory seeks to find most efficient ways of forming coalitions (teams) of agents, maximizing their gain and payoff, and also distributing gain in a fair way to keep coalitions stable by having all agents satisfied. We first propose a representation function for communities of web services based on their qualities and capacities. Using this function we can evaluate the $worth$ of each coalition of web services. Our community managers facing new membership requests, analyses whether the new coalition having the new set of web services will keep the community stable for $all$ web services already residing in the community or not. The community manager will reject the membership requests if it finds out new coalition can be unstable since there can be $any$ subset of web services gaining significantly more by deviating, colluding with new web services and joining other communities. Also web services facing communities, will try to join communities which maximizes their gain, if they find out they can deviate and agin more, they will form new coalitions with other communities. The cooperative game theory solution methods are combinatorial in nature, making them impractical in real world applications; however, using the concepts of group stability we have offered approximation algorithms running in polynomial time providing web services and community owners with applicable and near-optimal decision making mechanisms.

The rest of paper is organized as follows. Section \ref{s:preliminaries} describes architecture and parameters and some basic cooperative game theory concepts, section \ref{s:model} presents our solution model in three different scenarios. Section \ref{s:resutls} presents our experiments and results and finally section \ref{s:conclusion} concludes the paper and identified future work.

%In prevoius works \cite{Chun02user-centricperformance, mcknight00a, DBLP:journals/internet/BenatallahSD03, Khosravifar:2009:AIR:1586636.1586906, Rosario:2008:PQS:1512146.1512290}

%TODO Problems of \cite {DBLP:conf/IEEEscc/LimTMB12} having to cop and optimizing for with all weak web services reduces the optimal result\\

@@

We adopted a real world dataset \cite{DBLP:conf/smc/Al-MasriM09a} which has normalized and aggrigated each of these metrics to a real value between 0 and 1. Since tasks are not shared between web services and are distributed among all web serivices inside a community, one can expect tasks within a community to be performed by average quality of web service residing in that community. We use this average expected quality output, later in evaluating community $worth$ or $payoff$ function.

@@

\section{Problem Formulation and Modeling}\label{s:model}

@@

\emph{Proposition: } The folowwing statement is equivalent with convex condition of equation \ref{eq:convex}.
\begin{equation}\label{eq:convex_snow}
v(S \cup \left\{i\right\}) - v(S) \leq v (T \cup \left\{i\right\}) - v(T), \\ \forall S \subseteq T \subseteq N \backslash \left\{i\right\}, \forall i \in N; 
\end{equation}

@@

In coalition-formation games, formation of the coalitions is the most important aspect of in the game. These solutions focus on maximizing social welfare. For any coalition structure $\pi$, let $v_{cs}(\pi)$ denote the total worth $\sum_{C \in \pi}{v(C)}$ which we call it social welfare. The problems in this area deal with finding the maximum value for social welfare over all the possible coalition structures $\pi$. There are $centralized$ algorithms for this end, however these approaches are generally NP-complete. The reason is that number of all possible partitions of a set $N$ grows exponentially with the number of players in $n$, and the centralized algorithms need to iterate through all these partitions. 
%These algorithms \cite{DBLP:conf/ijcai/GrecoMPS11, DBLP:conf/ijcai/RahwanMJ11, RePEc:wpa:wuwpga:0110001} are centralized algorithms, where all the complexity. However these algorithms are more intractable than Core stable solutions and practical with some constraints in practice\cite{RePEc:wpa:wuwpga:0110001}. 
In our application, we prefer not having a centralized decision maker dictating all web services their decisions. Becuase of this plus the complexity issues we would prefer a distributed algorithm where each community master can be a decision maker. The aim is to find low complexity and distributed algorithms for forming coalitions\cite{DBLP:journals/igtr/AptW09,Dieckmann02dynamiccoalition,ray2007game}. The distributed merge-and-split algorithm in \cite{DBLP:journals/igtr/AptW09} suits our application very well. It keeps splitting and merging coalitions where it is preferred by all players inside those coalitions. 
This merge-and-split algorithm is designed to be adoptable for different problems. It requires two functions to be specified, first a preference relation or comparison relation for comparing two collections of different coalition partitions of same players and second function is to assess the stability condition of a partition. Having two partition sets of a subset of our players, namely $P = {P_1,...,P_k}$ and $Q = {Q_1,...N_l}$ one example would be to use the social welfare comarison $\sum^k_{i=1}v(P_i) > \sum^k_{i=1}v(N_l)$. For our scenario we used $Pareto order$ comparison which is an individual-value order appropriate for our self-interest agents (web services). In Pareto order, an allocation or partition $P$ is preferred over another $Q$ if at least one player gets more payoff in new allocation without other players getting hurt, or other players at least same payoff. ($p_i \geq q_i$ with at least one element $p_i > q_i$). As new web services are discovered and get ready to join communities, this algorithm keeps merging and splitting partitions based on the preference or comparison function provides and will terminate\cite{DBLP:journals/corr/abs-cs-0605132} as the partitions satisfy the stability condition or no further merging and splitting is found satisfying the comparison function. More details of this algorithm and analyses of generic solutions on coalition formation games are described in \cite{DBLP:journals/igtr/AptW09}.

@@

\section{Experimental Results}\label{s:resutls}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=3in]{task_done_opt.eps}
\includegraphics[width=3in]{task_qos_opt.eps}
\caption{Part (a): Cumulative number of tasks succesfully done. Part
(b): Average QoS of tasks performed.} \label{performanceall}
\end{figure}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=3in]{least_core.eps}
\caption{Smallest value of $\epsilon$ such that the $\epsilon-core$ is non-empty} \label{f_leastcore}
\end{figure}

@@

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=3in]{s2_task_done.eps}
\includegraphics[width=3in]{s2_task_qos.eps}
\caption{Part (a): Cumulative number of tasks succesfully done. Part
(b): Average QoS of tasks performed.} \label{performancemany}
\end{figure}

@@

\section{Conclusion}\label{s:conclusion}
